Overrelaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport

### 2017-11-28 PosterNIPS2017.tex
Last image of poster
Poster: Final details

			\item Converges \textbf{between 2 and 20 times faster} than SK, with almost the same cost per iteration, and other nice properties.
		\end{itemize}
	\end{minipage}
	\begin{minipage}{0.45\textwidth}
		\centering
		\begin{tcolorbox}[width=0.99\textwidth,/tcb/size=small, /tcb/top=2.0mm]
			\centering
			\begin{tabular}{c c c c}
				\includegraphics[height=3cm]{1d_interp_mu} &
				\includegraphics[height=3cm]{1d_interp_0_2} &
				\includegraphics[height=3cm]{1d_interp_2} &
				\includegraphics[height=3cm]{1d_interp_18} \\
				&
				$\epsilon = 5$ &
				$\epsilon = 0.5$ &
				$\epsilon = 0.02$
			\end{tabular}
			\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
		\end{tcolorbox}
	\end{minipage}
}

@@

\headerbox{Experiments}{name=expe,below=overrelaxation,column=1,span=2}{
	\begin{minipage}{\textwidth}
		We compare our algorithm to SK algorithm on two different settings of size $100 \times 100$.
		In both experimental settings, our algorithm wins by a factor above $20$ for small $\epsilon$.
	\end{minipage}
	\begin{minipage}{0.6\textwidth}
		\begin{tcolorbox}[size=small]
			\begin{minipage}{0.5\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{images/speedratio_image}\\
				(a) Squared euclidean norm on $[0,1]$. Marginals randomly generated, piecewise constant.
			\end{minipage}%
			\begin{minipage}{0.5\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{images/speedratio_ML}\\
				(b) Random cost, uniform marginals.
			\end{minipage}
		\captionof{figure}{Relative speed of our algorithm compared to Sinkhorn--Knopp, for various values of $\epsilon$}
		\end{tcolorbox}
	\end{minipage}
	\begin{minipage}{0.4\textwidth}
		\begin{tcolorbox}[size=small]
			\centering
			\includegraphics[width=0.89\textwidth]{typical_convergence} \\
			\captionof{figure}{Typical behavior of the algorithm: value of the Lyapunov function and of the overrelaxation parameters during the iterations.}
		\end{tcolorbox}
	\end{minipage}
}

@@

\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro,bottomaligned=expe}{
	Entropic regularization of optimal transport:
	\begin{equation*} \label{eq:problem}
	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
	\end{equation*}
	Marginal contraint sets $\Ccal_1$ and $\Ccal_2$ correspond to two discrete measures $\mu^1$ and $\mu^2$, and $\KL$ is the Kullback--Leibler divergence.
	\begin{gather*}
	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
	\quad\quad\quad
	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
	\end{gather*}
	\begin{minipage}{0.55\textwidth}
		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \textbf{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
		\begin{gather*}
		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0).
		\end{gather*}
		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is simply performed by \textbf{scaling the rows or columns} of the matrix.
		Operator $\oslash$ denotes element-wise division:
		\begin{align*}\label{scaling}
		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
		a &=  {\mu^1}\oslash{A_1 \gamma} \\
		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
		\end{align*}
	\end{minipage}
	\begin{minipage}{0.44\textwidth}
		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
			\centering
			\input{schema_a}
			\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
		\end{tcolorbox}
	\end{minipage}

>>>>

			\item Converges \textbf{between 2 and 20 times faster} than SK, with almost the same cost per iteration, and other nice properties.
		\end{itemize}
	\end{minipage}
	\begin{minipage}{0.45\textwidth}
		\centering
		\begin{tcolorbox}[width=0.99\textwidth,/tcb/size=small, /tcb/top=2.0mm]
			\centering
			\begin{tabular}{c c c c}
				\includegraphics[height=3cm]{1d_interp_mu} &
				\includegraphics[height=3cm]{1d_interp_0_2} &
				\includegraphics[height=3cm]{1d_interp_2} &
				\includegraphics[height=3cm]{1d_interp_18} \\
				&
				$\epsilon = 5$ &
				$\epsilon = 0.5$ &
				$\epsilon = 0.02$
			\end{tabular}
			\captionof{figure}{Regularized optimal transport maps become more and more precise for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
		\end{tcolorbox}
	\end{minipage}
}

@@

\headerbox{Experiments}{name=expe,below=overrelaxation,column=1,span=2}{
	\begin{minipage}{\textwidth}
		We compare our algorithm to SK algorithm on two different settings of size $100 \times 100$.
		In both experimental settings, our algorithm wins by a factor above $20$ for small $\epsilon$.
	\end{minipage}
	\begin{minipage}{0.6\textwidth}
		\begin{tcolorbox}[size=small]
			\begin{minipage}{0.5\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{images/speedratio_image}\\
				(a) Squared euclidean norm on $[0,1]$. Marginals randomly generated, piecewise constant.
			\end{minipage}%
			\begin{minipage}{0.5\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{images/speedratio_ML}\\
				(b) Random cost, uniform marginals.
			\end{minipage}
		\captionof{figure}{Relative speed of our algorithm compared to Sinkhorn--Knopp, for various values of $\epsilon$}
		\end{tcolorbox}
	\end{minipage}
	\begin{minipage}{0.4\textwidth}
		\begin{tcolorbox}[size=small]
			\centering
			\includegraphics[width=0.83\textwidth]{typical_convergence_2} \\
			\captionof{figure}{Typical run of the algorithm: value of the Lyapunov function and of the overrelaxation parameters during the iterations.}
		\end{tcolorbox}
	\end{minipage}
}

@@

\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro,bottomaligned=expe}{
	Entropic regularization of optimal transport:
	\begin{equation} \label{eq:problem}
	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
	\end{equation}
	Marginal contraint sets $\Ccal_1$ and $\Ccal_2$ correspond to two discrete measures $\mu^1$ and $\mu^2$, and $\KL$ is the Kullback--Leibler divergence.
	\begin{gather*}
	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
	\quad\quad\quad
	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
	\end{gather*}
	\begin{minipage}{0.55\textwidth}
		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \textbf{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
		\begin{gather*}
		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0).
		\end{gather*}
		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is simply performed by \textbf{scaling the rows or columns} of the matrix.
		Operator $\oslash$ denotes element-wise division:
		\begin{align*}\label{scaling}
		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
		a &=  {\mu^1}\oslash{A_1 \gamma} \\
		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
		\end{align*}
	\end{minipage}
	\begin{minipage}{0.44\textwidth}
		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
			\centering
			\input{schema_a}
			\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
		\end{tcolorbox}
	\end{minipage}
